---
title: "Community analysis"
output:
  html_document:
    keep_md: yes
editor_options:
  chunk_output_type: console
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# ipak function: install and load multiple R packages.
# check to see if packages are installed. Install them if they are not, 
#then load them into the R session.
# Source: https://gist.github.com/stevenworthington/3178163
ipak <- function(pkg){
new.pkg <- pkg[!(pkg %in% installed.packages()[,"Package"])]
if (length(new.pkg)) 
    install.packages(new.pkg, dependencies = TRUE)
sapply(pkg, require, character.only = TRUE)
}

#Loading packages
packages <- c("phyloseq","ggplot2","RColorBrewer","plyr","tidyverse",
              "knitr","magrittr","ape","vegan", "ggtree", 
              "cowplot", "ggrepel", "gtable", "gridExtra", "biom", "ampvis", "here")
ipak(packages)

#Standard error function
std <- function(x) sd(x)/sqrt(length(x))
CVcoeff <- function(x) (sd(x)/mean(x))*100
```

## Importing amplicon data

Read the `biom` files to do the analysis of the data limiting to the fields sampled in Michigan in 2011 and 2012.

```{r}
#File reading
Corn_COI_file <- here("data","COI_soil_otu_table_open.biom")
CornCOI <- import_biom(Corn_COI_file)

#Filtering samples to Corn fields
Corn_COI <- prune_samples((grepl('MICO', sample_names(CornCOI))), CornCOI)

#Rename tax ranks to actual names
colnames(tax_table(Corn_COI)) <- c("Phylum","Class","Order",
                                  "Family","Genus","Clade","Species","Kingdom")
```


## Community composition based on amplicon data

```{r, message=FALSE, warning=FALSE, fig.height=10, fig.width=12, fig.align="center"}
#Summarizing by field
Oom_phylo <- tax_glom(Corn_COI, taxrank = "Clade")
Oom_phylo <- merge_samples(Oom_phylo, "group")

#Pruning the top 20 taxa
most_taxa <- sort(taxa_sums(Oom_phylo), TRUE)[1:15]
Oom_phylo_top20 <- prune_taxa(names(most_taxa), Oom_phylo)
ntaxa(Oom_phylo_top20)

#Transform counts for plot
#Oom_phylo_state <- transform_sample_counts(Oom_phylo_state, 
#                                           function(x) 100 * x/sum(x))
State_Year <- psmelt(Oom_phylo_top20)

#Color scale
pal <- colorRampPalette(brewer.pal(12, "Paired"))

#Reorder factor
Clade_factor <- State_Year %>% group_by(Clade) %>% dplyr::summarise(sum(Abundance))
Clade_factor <- Clade_factor[order(-Clade_factor$`sum(Abundance)`),]
Clade_factor <- Clade_factor$Clade
State_Year$Clade <- factor(State_Year$Clade, levels = Clade_factor)
levels(State_Year$Clade)

data_state <- dplyr::select(State_Year, Sample, Abundance, Clade)
data_state <- data_state[with(data_state, order(Clade, as.numeric(Clade))),]

#Plot
(ByState <- ggplot(data = data_state, aes(Sample, Abundance, fill = Clade)) +
  geom_bar(stat = "identity", position = position_fill()) + coord_flip() +
  scale_fill_manual(values = pal(20)) + 
  theme_gray() + 
  theme(text = element_text(size = 15)))
```


## Culture data: isolation frequency

Initial analysis of culture data and plotting frequency of isolation per species

```{r, fig.align="center", fig.width=8, fig.height=14}

#Importing otu like matrix, tax table and sample data for culture collection
culture_otu <- read.table(here("data","otu_corn_culture.txt")) %>% as.matrix()

culture_tax <- read.table(here("data","tax_corn_culture.txt")) %>% as.matrix() 

culture_data <- read.table(here("data","data_corn_culture.txt"), sep = "\t", header = TRUE, row.names = "group")

otu_corn_ct <- otu_table(culture_otu, taxa_are_rows = TRUE)
tax_corn_ct <- tax_table(culture_tax)
data_corn_ct <- sample_data(culture_data)

#Creating phyloseq object for culture collection
corn_ct <- phyloseq(otu_corn_ct, tax_corn_ct, data_corn_ct)

#Species recovery by isolation
corn_ct_sp <- psmelt(corn_ct)
corn_ct2 <- corn_ct_sp %>% group_by(Species, Year) %>% summarise(Abd = sum(Abundance))
corn_ct2$Year <- as.factor(corn_ct2$Year)
corn_ct2$Species <- gsub("_"," ", corn_ct2$Species)

(ct_bar_plot <- ggplot(corn_ct2) + 
  geom_bar(aes(x = reorder(Species, Abd, mean), y = Abd, fill = Year), stat = "identity") + 
  labs(x ="Species", y = "Frequency") +
  coord_flip() +
  scale_fill_manual(values = c("#7fbf7b", "#3288bd")) +
  theme_gray() +
  theme(text = element_text(size = 15),
        axis.text.y = element_text(face = "italic")))
```

## Importing culture data and merging data with amplicon phyloseq object

```{r, warning=FALSE}
#merging amplicon dataset by field
Oom_group <- merge_samples(Corn_COI, "group")
Oom_group <- subset_taxa(Oom_group, Clade != "NA")

#adding info to phyloseq files
sample_data(corn_ct)$type <- "Culture"
sample_data(Oom_group)$type <- "COI amplicon"
sample_names(corn_ct) <- paste(sample_names(corn_ct),"_ct",sep = "")
sample_names(Oom_group) <- paste(sample_names(Oom_group),"_COI",sep = "")

#transform abundances
corn_ct2 <- transform_sample_counts(corn_ct, function(OTU) (OTU/sum(OTU)))
Oom_group2 <- transform_sample_counts(Oom_group, function(OTU) (OTU/sum(OTU)))

#merge and transform abundances
ct_COI_corn <- merge_phyloseq(Oom_group, corn_ct)
ct_COI_corn2 <- merge_phyloseq(Oom_group2, corn_ct2)
```

### Comparing culture data with amplicon data

```{r, warning=FALSE, fig.align="center", fig.width=10, fig.height=12}
#Creating data frame with abundance data
ct_COI_corn.psmelt <- psmelt(ct_COI_corn)

ct_COI_corn_clade <- ct_COI_corn.psmelt %>% 
  group_by(Clade, type) %>% 
  summarise(abd = sum(Abundance)) %>%
  group_by(type) %>%
  mutate(total = sum(abd)) %>%
  mutate(Rel.abundance = (abd/total))


ct_COI_corn_clade <-  ct_COI_corn_clade %>%
  dplyr::select(Clade,type,Rel.abundance) %>% 
  spread(type, Rel.abundance) %>% 
  top_n(n =15, wt = `COI amplicon`) %>%
  gather(type, Rel.abundance, 2:3)
ct_COI_corn_clade$Clade <- gsub("_"," ", ct_COI_corn_clade$Clade)

(clade.graph <- ggplot(ct_COI_corn_clade) + 
  geom_bar(aes(x = reorder(Clade, Rel.abundance, mean), y = Rel.abundance), stat = "identity") + 
  facet_grid(. ~ type, scales = "free_y") +
  ylab("Relative abundance") +
  xlab("Genus/Clade") +
  coord_flip() + 
  theme_gray(base_size = 16) + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
        axis.text.y = element_text(size = 16),
        strip.text = element_text(face = "bold", size = 20)))


ct_COI_corn_genus <- psmelt(ct_COI_corn) %>% 
  group_by(Species, type) %>% 
  summarise(abd = sum(Abundance)) %>%
  group_by(type) %>%
  mutate(total = sum(abd)) %>%
  mutate(Rel.abundance = (abd/total))

ct_COI_corn_genus <-  ct_COI_corn_genus %>%
  dplyr::select(Species,type,Rel.abundance) %>% 
  spread(type, Rel.abundance) %>% 
  top_n(n=12, wt = `COI amplicon`) %>%
  gather(type, Rel.abundance, 2:3)
ct_COI_corn_genus$Species <- gsub("_"," ", ct_COI_corn_genus$Species)

(species.graph <- ggplot(ct_COI_corn_genus) + 
  geom_bar(aes(x = reorder(Species, Rel.abundance, mean), 
               y = Rel.abundance), stat = "identity") + 
  facet_grid(. ~ type, scales = "free_y") +
  xlab("Top 12 Species") +
  ylab("Relative abundance") +
  coord_flip() + 
  theme_gray(base_size = 16) + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
        axis.text.y = element_text(size = 18),
        strip.text = element_text(face = "bold", size = 20)))
```


## Alpha diversity for amplicon data

```{r, warning=FALSE}
#Plotting richness
alpha.corn <- plot_richness(Corn_COI, "group", 
              measures = c("InvSimpson","Shannon","Chao1"), color = "Year")

alpha.corn + geom_boxplot() + xlab("Field") + theme_gray() +
  theme(axis.text.x = element_text(angle = 90))
```

__Richness table__

```{r}
#Table richness
Tb_richness <- estimate_richness(Corn_COI, 
                                 split = TRUE, 
                                 c("Observed", "Shannon", "Simpson")) %>% 
  rownames_to_column(var = "sample") %>%
  mutate(Evenness = Shannon/log(Observed))

samp_size <- colSums(otu_table(Corn_COI))
samp_size <- data.frame(samp_size) %>% rownames_to_column(var = "sample")

smp_state <- data.frame(sample_data(Corn_COI)[,71:72]) %>% 
  rownames_to_column(var = "sample")

Tb_richness_final <- left_join(Tb_richness, samp_size, by = "sample") %>%
  left_join(smp_state, by = "sample") %>%
  filter(Observed > 1) %>%
  ddply(c("group"), summarise, 
        N = length(sample),
        #OTUs = sum(samp_size),
        mean.Observed = mean(Observed),
        sd.Observed = sd(Observed, na.rm = TRUE),
        mean.Shannon = mean(Shannon),
        sd.Shannon = sd(Shannon, na.rm = TRUE),
        mean.Simpson = mean(Simpson),
        sd.Simpson = sd(Simpson, na.rm = TRUE),
        mean.Evenness = mean(Shannon/log(Observed)),
        sd.Evenness = sd(Shannon/log(Observed), na.rm = TRUE))

kable(Tb_richness_final, digits = 3, format = "markdown")
```

### ANOSIM

```{r}
Yr_grp <- get_variable(Corn_COI, "Year")
(Corn_yr_anosim <- anosim(phyloseq::distance(Corn_COI, "bray"), Yr_grp))

field_grp <- get_variable(Corn_COI, "group")
(Corn_grp_anosim <- anosim(phyloseq::distance(Corn_COI, "bray"), field_grp))
```


## Beta diversity

### ADONIS

```{r}
df <- as(sample_data(Corn_COI), "data.frame")
d <- phyloseq::distance(Corn_COI, "bray")

Corn_adonis <- adonis(d ~ Year, df)

kable(Corn_adonis$aov.tab, digits = 3, 
      caption = "__Table .__ Comparison of community structure (beta diversity)\
      using Bray-curtis distance by year.", format = "markdown")
```


### Ordination analysis for culture and amplicon

```{r Ordination_plot}

##Colors
colors2 <- c("#77C7C6",
"#C37F3B",
"#869BCF",
"#7DD54E",
"#C67BB7",
"#CDC84A",
"#CC6569",
"#83D693",
"#7A7678",
"#698547",
"#D3C1A7",
"#6ca556")

##Year as factor
sample_data(ct_COI_corn)$Year <- as.factor(sample_data(ct_COI_corn)$Year)

##Beta diversity culture and amplicon
Oom_biom_ord <- ordinate(ct_COI_corn, "PCoA", "bray")
ord_plot <- plot_ordination(ct_COI_corn, Oom_biom_ord, color = "Year", shape = "type")
(ord_plot.COI <- ord_plot + geom_point(size = 5, alpha = 0.7) + 
  scale_colour_manual(values = colors2) + labs(title = "coxI") + 
  theme_gray(base_size = 18) + labs(color = "Year", shape = "Type")) 

```


## Beta diversity for amplicon data only

```{r}
#Extracting the top 100 taxa
most_taxa <- sort(taxa_sums(Corn_COI), TRUE)[1:100]
Corn_COI_top100 <- prune_taxa(names(most_taxa), Corn_COI)

#Ordination analysis
Oom_biom_ord2 <- ordinate(Corn_COI_top100, "PCoA", "bray")
ord_plot2 <- plot_ordination(Corn_COI_top100, Oom_biom_ord2, color = "group", shape = "Year")
(ord_plot.COI <- ord_plot2 + geom_point(size = 5, alpha = 0.7) + 
  scale_colour_manual(values = colors2) + labs(title = "PCoA COI") +
  theme_gray(base_size = 18) + labs(color = "group", shape = "Year") +
  theme(legend.text = element_text(size = 20),
        legend.key.size = unit(0.9, "cm")))
```

### Environmental/Edaphic factor analysis

```{r envfit_analysis, warning=FALSE}
## Environment fit analysis
bray.pcoa <- ordinate(Corn_COI_top100, method = "PCoA", "bray")
env <- sample_data(Corn_COI_top100)
env <- env[,c(1,6,7,8,10,11,15,16,17,18,69,73)]
env[,1:12] <- sapply(env[,1:12], as.numeric)
str(env)


Oom_env <- envfit(bray.pcoa$vectors, env, permutations = 999, na.rm = TRUE)

fit_data <- as.data.frame(scores(Oom_env, display = "vectors")) %>%
  add_rownames(var = "Env.var") %>%
  bind_cols(data.frame(Oom_env$vectors$r, Oom_env$vectors$pvals)) %>%
  #rename(R2 = Oom_env.vectors.r, P.value = Oom_env.vectors.pvals) %>%
  arrange(Oom_env.vectors.pvals)
  
## Supplementary material version

kable(fit_data, digits = 3, caption = "Significance and correlation\
of vectors fitted into PCoA ordination of oomycete communities associated with\
Corn seedlings", format = "markdown")

```

### Results ordination and environmental data

```{r Ord_envfit_plot, warning=FALSE,fig.align='center', fig.width=10, fig.height=6}
## Vectors for plot
fit_reduced <- fit_data[fit_data$Oom_env.vectors.pvals < 0.05,] 

fit_plot <- as.data.frame(scores(Oom_env, display = "vectors")) %>%
  add_rownames(var = "Env.var") %>%
  inner_join(fit_reduced, by = "Env.var") %>%
  arrange(Oom_env.vectors.pvals) 

fit_plot$Env.var2 <- c("Clay (%)","Organic matter (%)", "Water content",
                       "Soil pH","Bulk density (g/cm^3)",
                       "CEC (meq/100g soil)",
                       "Avail. water capacity (cm water/cm soil)")

ord_plot.data <- plot_ordination(Corn_COI_top100, Oom_biom_ord2, 
                            color = "group", shape = "Year", justDF = TRUE)

(ord.plot.env <- ggplot(data = ord_plot.data, aes(x = Axis.1, y = Axis.2)) + 
  geom_point(aes(color=group, shape=Year), size = 6, alpha = 0.9) + 
  #scale_color_brewer(type = "div", palette ="Spectral") +
  labs(color = "group", shape = "Year", x = "PCoA 1 [18.2%]", y = "PCoA 2 [10.4%]") +
  scale_colour_manual(values = colors2) +
  geom_segment(data = fit_plot, aes(x = 0, xend = Axis.1.x, y = 0, yend = Axis.2.x), 
               arrow = arrow(length = unit(0.1,"cm")), color = "black", size = 1) + 
  geom_label_repel(data = fit_plot, aes(x = Axis.1.x, y = Axis.2.x, label = Env.var2), 
            size = 5, force = 1) + #facet_wrap(~Year) +
  theme_gray(base_size = 18) + theme(legend.title = element_text(size = 22),
                                     legend.text = element_text(size = 20),
                                     legend.key.size = unit(0.8, "cm")))
```


## Relative abundance of top 20 species

```{r, warning=FALSE, fig.align="center", fig.width=12, fig.height=14}
Corn_COI.f <- subset_taxa(Corn_COI, !is.na(Clade)) 
amp_heatmap(data = Corn_COI.f,
            group = "group",
            tax.show = 20,
            tax.aggregate = "Species",
            tax.add = "Clade")
```

